---
title: "Florida Schools Data Visualization"
output: html_notebook
---

```{r}
# init session
library(tidyverse); library(data.table);
library(ggplot2); library(gganimate); library(ggthemes); library(ggmap); library(ggrepel); library(gifski)
options(scipen=20)
df <- fread("sch_include_03_01_21.csv", integer64="numeric")
```
 
## Emergence of charter schools

Create a static plot of charter school emergence in Florida.
```{r fig.width=8,fig.height=6}
# get backdrop from stamen maps
florida_background <- get_stamenmap( bbox = c(left = -88, bottom = 24, right = -79.9, top = 31.2), zoom = 6, color='color', maptype = "terrain-background")

# create static plot
plot <- ggmap(florida_background, darken=c(0.5, 'white')) + 
  geom_point(data=df[distinct_charter == T], mapping=aes(x=gm_lon, y=gm_lat, size=k12_enrollment), alpha=0.2, color="red") +
  ggtitle("Emergence Of Charter Schools In FLorida") +
  labs(title = "Emergence Of Charter Schools In Florida", x="Longitude", y="Latitude", size="Enrollment:", subtitle="Year: xxxx")
plot
```

Create an animated plot from the previous static plot.
```{r}
# create animated plot w/ live labs
plot.animate <- plot +
  transition_time(year) +
  labs(subtitle = "Year: {frame_time}", 
       size="Enrollment:") +
  theme(plot.title=element_text(size=24), plot.subtitle=element_text(size=18), legend.title=element_text(size=18), 
        legend.text=element_text(size=14))

# Total Enrollment: {df$k12_enrollment[df$year==frame_time & df$distinct_charter == T] %>% sum() %>% format(big.mark=',')}

# produce animation
animate(plot.animate, height=600, width=700, nframes=24, fps=8, end_pause=4) #%>% anim_save(animation=., filename = "img/emergenceOfCharterSchools.gif")
```

## Measure of competition/presence

Initialize best-fit competition measure.
```{r}
# running more than once will cause extra columns in df, so don't do that

# faster computation with sparse matrices
library(Matrix)

# read distance matrix
dm <- fread("src/distance_matrix.csv", integer64="numeric") %>% as.matrix

# use best hyperparameter fit
dmax_set = c(4); a_set = c(0.3);

# all grades to iterate through and average
grades <- paste0(c("k",paste0("grade",1:12)),"_enr")

y_enr = list(); y_cenr = list()
for (y in 1999:2018) {
    # create a list of enrollments by year. row is location ID, column is grade
    y_enr[[y]] <- (merge(data.table(location_id=1:nrow(dm)),df[year == y, lapply(.SD,function(x) sum(x,na.rm=T)), by=location_id, .SDcols = grades], by="location_id", all.x=T) %>% select(2:14) %>% replace(.,is.na(.),0) %>% as.matrix)
    # & create a list of charter enrollments by year. row is location ID, column is grade
    y_cenr[[y]] <- (merge(data.table(location_id=1:nrow(dm)),df[year == y & charter=="yes", lapply(.SD,function(x) sum(x,na.rm=T)), by=location_id, .SDcols = grades], by="location_id", all.x=T) %>% select(2:14) %>% replace(.,is.na(.),0) %>% as.matrix)
}


# LOOP ----
# iterate through all values of max distance
for (dmax in dmax_set) { 
  # where distance >= distance max, set the weight to 0
  dmf <- (dm/dmax); dmf[dmf > 1] <- 1
  
  # iterate through all curve-shaping values
  for (a in a_set) {
    paste0("\na=",a,"; dmax=",dmax) %>% cat
    # apply the curving function to the distance matrix
    dmf <- Matrix(1-(dmf^a),sparse=T)
        
    # iterate through all years
    for (y in 1999:2018) {
      
      # STUDENT PRESENCE proximate to location I, weighted by distance curving function
      sp <- data.table(as.matrix(dmf %*% y_enr[[y]])); setnames(sp,paste0('sp_',grades))
      # CHARTER STUDENT PRESENCE proximate to location I, weighted by distance curving function
      csp <- data.table(as.matrix(dmf %*% y_cenr[[y]])); setnames(csp,paste0('csp_',grades))
      # add location_id and year columns so that a join with df is possible
      # if this throws an error it means your dm and 
      sp[,`:=`(location_id = 1:nrow(dm), year=y)]; csp[,`:=`(location_id = 1:nrow(dm), year=y)]
      
      # data.table of location ids and 
      if (y==1999) t_lids <- merge(sp,csp,on=.(location_id,year))
      else t_lids <- rbind(t_lids, merge(sp,csp,on=.(location_id,year)))
    }
  
    # generate a column representing charter student presence given dmax and a
    t_m <- merge(select(df,location_id,year,ncessch,k12_enrollment,grades),t_lids,by=c("location_id","year"))
    t_m[[paste0("csp","_a",a,"_dmax",dmax)]] <- (rowSums((as.matrix(select(t_m, paste0("csp_",grades))) / as.matrix(select(t_m, paste0("sp_",grades,""))))  *  as.matrix(select(t_m,grades))) / as.matrix(t_m$k12_enrollment))
    df <- merge(df,select(t_m,year,ncessch,paste0("csp","_a",a,"_dmax",dmax)),by=c("ncessch","year"))
  }
}
```

Let's focus on a specific school to get an idea of how the competition/presence measures are calculated. The school we're focusing on is a TPS in Lakeland called Oscar J. Pope Elementary. ncessch=120159001783, location_id=3698, 
```{r fig.width=7,fig.height=6}
# get backdrop from stamen maps
florida_background <- get_stamenmap( bbox = c(left = -81.98, bottom = 27.94, right = -81.83, top = 28.08), zoom = 12, color='color', maptype = "terrain-lines")

# find all the locations within 8 miles of the one we're looking at
proximateSchools = which(dm[3698,] < 5) %>% unname

# create static plot over years y
y = c(2018)
plot <- ggmap(florida_background, darken=c(0.5, 'white')) + 
  geom_point(data=df[year%in%y & location_id%in%proximateSchools,# & location_id!=3698, 
                     .(k12_enrollment=sum(k12_enrollment), gm_lon=first(gm_lon), gm_lat=first(gm_lat), distinct_charter=first(distinct_charter)), 
                     by=.(year, location_id)], 
            mapping=aes(x=gm_lon, y=gm_lat, size=k12_enrollment, color=factor(distinct_charter, labels=c("TPS", "Charter")), alpha=round(1-(dm[3698, location_id]/4)^0.3, 2))) +
  
  geom_text_repel(data=df[year%in%y & location_id==3698], 
            mapping=aes(x=gm_lon, y=gm_lat, label="Oscar J. Pope Elementary"), nudge_y=0.004, nudge_x=0.04, segment.size=0.17, segment.color='black') +
  
  geom_text_repel(data=df[year%in%y & location_id==3759], 
            mapping=aes(x=gm_lon, y=gm_lat, label="South McKeel Academy (K-7)"), hjust=0, vjust=0, nudge_y=0.012, nudge_x=0.025, segment.size=0.2, segment.color='black') +
  
  scale_size(range = c(4, 12)) +
  labs(title = "Charter Competition for Oscar J. Pope (2018)", x="Longitude", y="Latitude", size="Enrollment", color='School type', alpha="Presence in \nweighted average") +
  theme(axis.ticks = element_blank(), axis.text = element_blank(), axis.title = element_blank()) +
  scale_color_manual(values=c("dodgerblue3", "orangered3")) +
  theme(plot.title=element_text(size=18), plot.subtitle=element_text(size=12), legend.title=element_text(size=12), 
        legend.text=element_text(size=11))

# print/save plot
plot #%>% ggsave(filename="img/competitionMeasureExample.jpg")
```


